A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints

### 2021-05-24 FMCAD2021/1_intro.tex
sec1,4: complexity

Aiming at introducing Point's procedure to computer science researchers, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science researchers (Section~\ref{sec-dec}). This can be seen as another contribution of this paper.
%
\wuhao{We also analyze the deterministic upper bound of the  complexity is triply exponential for the first time (Appendix~\ref{app:cpx}). Since the best known upper bound for Cooper's elimination algorithm is also triply exponential, it may not be easy to improve the deterministic upper bound. }
Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).

>>>>

Aiming at introducing Point's procedure to computer science researchers, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science researchers (Section~\ref{sec-dec}). This can be seen as another contribution of this paper.
%
We also analyze the deterministic upper bound of the  complexity is triply exponential for the first time (Appendix~\ref{app:cpx}). \wuhao{Since the best known upper bound for Cooper's elimination algorithm is also triply exponential, it may not be easy to improve the bound. }
Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).

### 2021-05-24 FMCAD2021/paper.tex
sec1,4: complexity

\begin{abstract}
    String constraint solving is the core of various testing and verification approaches for scripting languages. 
    Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
    As string-integer conversion is an important function appearing in almost all scripting languages, Abdulla et al. extended the flattening approach to this function recently.
    However, their approach supports only a special flattening pattern and leaves the support of the general flat regular constraints as an open problem.
    In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion. The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986. \wuhao{We analyze the complexity of the algorithm and give a triply exponential upper bound both in time and space.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
    We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.
    The experimental results show that our implementation outperforms the state-of-the-art solvers.

>>>>

\begin{abstract}
    String constraint solving is the core of various testing and verification approaches for scripting languages. 
    Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
    As string-integer conversion is an important function appearing in almost all scripting languages, Abdulla et al. extended the flattening approach to this function recently.
    However, their approach supports only a special flattening pattern and leaves the support of the general flat regular constraints as an open problem.
    In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion. The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986. \wuhao{We analyze the complexity of the algorithm and give a triply exponential upper bound both in deterministic time and space.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
    We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.
    The experimental results show that our implementation outperforms the state-of-the-art solvers.
